### 2021-03-04 paper/main.tex
Added software paper

\documentclass[12pt, a4paper]{article}

% Set the document margins for [A4 paper; and, left, right, top, and bottom margins; the text to footer gap; and include the footer in the main body of the text (so page numbers leave space equal to the bottom value just specified)]
\usepackage[a4paper, left=2.5cm, right=2cm, top=2cm, bottom=2cm, footskip=1.5cm, includefoot]{geometry}

\usepackage{natbib}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{hyperref} % to insert a web link
\usepackage{amsfonts} % to get blackboard font

\begin{document}

\title{compGeometeR: an R package for computational geometry}

\author{Thomas R. Etherington  \\
	Manaaki Whenua -- Landcare Research  \\
	\and 
	O. Pascal Omondiagbe \\
	Manaaki Whenua -- Landcare Research \\
	}

\date{\today}

\maketitle

\begin{abstract}

Computational geometry algorithms and data structures are widely applied across numerous scientific domains, and there a variety of R packages that apply computational geometry.  However, these packages often work in specific numbers of dimensions, have incompatible data structures, and include additional non-computational geometry functionality that can be domain specific.  Our objective in developing the compGeometeR package is to produce a systematic implementation of the most commonly used computational geometry algorithms so that they can be easily integrated into domain specific scientific workflows.  We briefly explain the algorithms available in compGeometeR, and identify priorities for future development.

\begin{center}
\textbf{Keywords}: R, alpha complex, alpha shape, convex hull, convex layers, Delaunay triangulation, digital, discrete.
\end{center}

\end{abstract}

\section{Introduction}

Computational geometry seeks to develop algorithms and data structures for geometric objects, and the results are widely applied across numerous scientific domains \citep{de-berg-2008}.  Given the importance of computational geometry it is no surprise that within the R computing ecosystem there are a variety of options for applying various computational geometry algorithms in a variety of dimensions.  However, these packages often work in specific numbers of dimensions, have incompatible data structures, and include additional non-computational geometry functionality that can be domain specific.

Our objective in developing the compGeometeR package is to produce a systematic implementation of some of the most commonly used computational geometry algorithms across various of branches of geometry so that they can be easily integrated into domain specific scientific workflows.  The functionality of compGeometeR is entirely dependent on computational foundations provided by the Quickhull software \citep{barber-1996}.

\section{Algorithms}

There are a variety of computational geometry algorithms available in compGeometeR, some of which are replicated in other R packages.  By documenting these differences (Table \ref{tab:r-options}) we hope to help potential users to assess if compGeometeR is the best option for their needs as there may be other options that are more suitable.

% The following table uses formatting fucntions of the booktabs package: https://texdoc.org/serve/booktabs/0
\begin{table}
\small
\centering
\caption{Computational geometry software options in R.  For each computational geometry function, the R software (and version) that can apply this algorithm is noted by reference to the number of dimensions in which the algorithm will work.}
\begin{tabular}{l c c c c c c c c}
  \toprule

   & \rotatebox{90}{compGeometeR (1.0.0)}
   & \rotatebox{90}{R (3.5.3)}
   & \rotatebox{90}{alphahull (2.2)}
   & \rotatebox{90}{alphashape3d (1.3.1)}
   & \rotatebox{90}{deldir (0.1-25)}
   & \rotatebox{90}{geometry (0.4.5)}
   & \rotatebox{90}{spatstat (1.63-3)}
   & \rotatebox{90}{tripack (1.3-9.1)} \\

  \cmidrule{1-9} 
  Discrete geometry		&     &   &     &   &   &     &   &   \\
  \cmidrule{1-9} 
  Alpha complex				& $n$ &   &     &   &   &     &   &   \\
  Alpha shape				&     &   &  2  & 3 &   &     &   &   \\
  Convex hull  				& $n$ & 2 &     &   &   & $n$ & 2 & 2 \\
  Convex layers				& $n$ &   &     &   &   &     &   &   \\
  Delaunay triangulation	& $n$ &   &  2  &   & 2 & $n$ & 2 & 2 \\
  Voronoi diagram			&     &   &  2  &   & 2 & $n$ & 2 & 2 \\

  \cmidrule{1-9} 
  Digital geometry			&     &   &     &   &   &     &   &   \\
  \cmidrule{1-9} 
  Alpha complex				& $n$ &   &     &   &   &     &   &   \\
  Alpha shape				& $n$ &   &     &   &   &     &   &   \\
  Convex hull  				& $n$ &   &     &   &   &     &   &   \\
  Delaunay triangulation	& $n$ &   &     &   &   &     &   &   \\
  \bottomrule
\end{tabular}
\label{tab:r-options}
\end{table}

\subsection{Discrete geometry}
% see: https://en.wikipedia.org/wiki/Discrete_geometry

Discrete geometry studies the shapes and interactions between sets of discrete objects such as points, lines, circles, and polygons in 2-dimensional space.  All of the discrete computioanl geomtery algorithms in compGeometeR take as an input a set of points $P$ in $n$-dimensional Eucliden space $\mathbb{R}^n$.

A convex hull defines the smallest subset of $\mathbb{R}^n$ that contains $P$ and for which the subset is convex so any two points within the convex hull can be connected by a straight line also contained by the convex hull (Figure \ref{fig:euclidean-algorithms}a).

Convex layers were first presented by \cite{huber-1972} and \cite{barnett-1976} who both gave unreferenced credit for this idea to Tukey.  Convex layers are a nested sequence of convex hulls produced by repeating the process of constructing a convex hull for $P$ and then removing the points forming the vertices of the convex hull from $P$ before producing the next convex layer.  The first convex layer is equivalent to the convex hull, with each successive convex layer representing ever smaller region of space (Figure \ref{fig:euclidean-algorithms}b).

The Delaunay triangulation \citep{delaunay-1934} produces a set of simplices (triangles in 2-dimensions or tetrahedrons in 3-dimensions) for which no point in $P$ is inside the the circumhypersphere (a circle in 2-dimensions or a sphere in 3-dimensions) of each simplex (Figure \ref{fig:euclidean-algorithms}c).

The alpha complex \citep{edelsbrunner-1994} is a subset of the Delaunay triangulation that contains only those simplices for whose circumhypersphere radius is smaller than a specified $\alpha$ parameter value ranging $0 < \alpha < \infty$ (Figure \ref{fig:euclidean-algorithms}d).

\begin{figure}[ht]
\centering
\includegraphics{figures/figure-1/figure-1.png}
\caption{Two-dimensional examples of discrete geometry algorithms currently available in compGeometeR. (a) Convex hull, (b) convex layers, (c) Delaunay triangulation, and (d) alpha complex.}
\label{fig:euclidean-algorithms}
\end{figure}

\subsection{Digital geometry}
% see: https://en.wikipedia.org/wiki/Digital_geometry

Digital geometry is a branch of geometry that studies the geometric properties of a grid (or lattice) of points in Euclidean space, and usually involves the digitisation of discrete lines and regions \citep{rosenfeld-1989}.

\section{Future work}

compGeometeR is very much a work in progress, and while there is sufficient functionality for it to be of wide use in the computational sciences, there is some functionality that has been identified as being particularly useful for future development.

\subsection{Geometric graph theory}
% see: https://en.wikipedia.org/wiki/Geometric_graph_theory

Geometric graph theory...

There are a variety of useful graph structures that can be produced based on geometric principles.  For example, the Delaunay triangulation \citep{delaunay-1934} can also be represented as a graph structure, and for which there are several useful subgraphs such as the Gabriel graph \citep{gabriel-1969}, Urquhart graph \citep{urquhart-1980}, and relative neighbourhood graph \citep{toussaint-1980}.

\subsection{Spherical and ellipsoidal geometry}
% see: https://en.wikipedia.org/wiki/Spherical_geometry
% see: https://en.wikipedia.org/wiki/Elliptic_geometry

All the algorithms in compGeometeR work in Cartesian space.

\subsection{Fuzzy geometry}

In some contexts it will be important to recognise that there is uncertainty in the position of points in space and the parameters defining a given algorithm.  In such situations it may be helpful to adopt fuzzy geometry \citep{rosenfeld-1998} view in which, foe example, memberships of points in space are not crisp being either 0 or 1, but rather are fuzzy on a scale from 0 to 1.  Creating fuzzy versions of the compGeometeR algorithms could result in very useful functionality where quantifying uncertainty of any computational geometry is important.  The existing digital geometric algorithms could be easily extended to fuzzy forms as all that is required is to

\section{Code availability}

compGeometeR is open source software made available under the GPL license. Installation
instructions can be found at the GitHub repository \url{https://github.com/manaakiwhenua/compGeometeR}.

\section{Acknowledgements}

This research was funded by the New Zealand Ministry of Business, Innovation and Employment via the Beyond Myrtle Rust (\#C09X1806) and Winning Against Wildings research programmes, with additional internal investment by Manaaki Whenua -- Landcare Research

\bibliographystyle{landscapeecol}
\bibliography{references}

\end{document}

### 2021-03-04 paper/figures/figure-1/figure-1.R

@@ -1,0 +1,69 @@
+#-------------------------------------------------------------------------------
+
+# R version 3.5.3 (2019-03-11) -- "Great Truth"
+# Platform: x86_64-w64-mingw32/x64 (64-bit)
+
+library(compGeometeR) # version 1.0
+
+#-------------------------------------------------------------------------------
+
+set.seed(2) # to reproduce figure exactly
+x = rgamma(n = 20, shape = 3, scale = 2)
+y = rnorm(n = 20, mean = 10, sd = 2)
+p = cbind(x, y)
+
+#-------------------------------------------------------------------------------
+
+# Function to sort vertices into a circular order
+circular_vertex_sort <- function(vertices) {
+  midpoint = colMeans(vertices)
+  angles = atan2(vertices[,1] - midpoint[1], vertices[,2] - midpoint[2])
+  angles[angles < 0] = angles[angles < 0] + 2 * pi
+  ch_vertices_order = sort(angles, index.return=TRUE)$ix
+  vertices[ch_vertices_order,]
+}
+
+#-------------------------------------------------------------------------------
+
+png("figure-1.png", width = 10, height = 10, res=300, units = "cm", pointsize = 10)
+par(mfrow=c(2,2), oma=c(0,0,0,0), mar=c(0.5,0.5,1.5,0.5))
+
+# Convex hull
+plot(x, y, yaxt="n", xaxt="n", xlab="", ylab = "", pch=16, cex=0.75)
+ch = convex_hull(p)
+circ_vertices = circular_vertex_sort(ch$hull_vertices)
+polygon(circ_vertices, col="lightgrey", border="darkgrey")
+points(p, pch=16, cex=0.75)
+mtext("(a)", side = 3, line=0.25)
+
+# Convex layers
+plot(x, y, yaxt="n", xaxt="n", xlab="", ylab = "", pch=16, cex=0.75)
+for (i in seq(4)) {
+  cl = convex_layer(p, layer = i)
+  circ_vertices = circular_vertex_sort(cl$hull_vertices)
+  polygon(circ_vertices, col="lightgrey", border="darkgrey")
+}
+points(p, pch=16, cex=0.75)
+mtext("(b)", side = 3, line=0.25)
+
+# Delaunay triangulation
+plot(x, y, yaxt="n", xaxt="n", xlab="", ylab = "", pch=16, cex=0.75)
+dt = delaunay(p)
[…]

### 2021-03-04 R/grid-coordinates.R
Updated grid-coordinates to accpet different spacings in each dimension

@@ -7,32 +7,36 @@
 #' dimension.
 #' @param maxs Vector of length \code{n} listing the point space maximum for each
 #' dimension.
-#' @param spacing The spacings between points in all dimensions.
+#' @param spacings Vector of length \code{n} listing the spacings between points 
+#' in all dimensions.
 #'
 #' @return A matrix with \code{n} columns - and potentially lots of rows!
 #'
 #' @examples
 #' # Point space grid coordinates usage
-#' xy = grid_coordinates(mins=c(0,0), maxs=c(10,15), spacing=1)
+#' xy = grid_coordinates(mins=c(0,0), maxs=c(10,15), spacings=c(1,1))
 #' 
 #' @export
-grid_coordinates <- function(mins, maxs, spacing) {
+grid_coordinates <- function(mins, maxs, spacings) {
   
   # Check input data
   if (length(mins) != length(maxs)) {
     stop("Length of mins and maxs differ")
   }
+  if (length(mins) != length(spacings)) {
+    stop("Length of spacings differs from length of mins and maxs")
+  }
   if (FALSE %in% (mins < maxs)) {
     stop("Maximums not greater than minimums in all dimensions")
   }
-  if (spacing <= 0) {
-    stop("Spacing must be greater than zero")
+  if (TRUE %in% (spacings <= 0)) {
+    stop("All spacings must be greater than zero")
   }
   
   # Create list of coordinate locations for each dimension
   dimCoords = list()
   for (n in seq(1, length(mins))) {
-    dimCoords[[n]] <- seq(mins[n], maxs[n], spacing)
+    dimCoords[[n]] <- seq(mins[n], maxs[n], spacings[n])
   }
   # Create all combinations of coordinates across all dimension
   return(expand.grid(dimCoords))

### 2021-03-05 paper/main.tex
Updated paper

@@
   Convex hull  				& $n$ &   &     &   &   &     &   &   \\
-  Delaunay triangulation	& $n$ &   &     &   &   &     &   &   \\
   \bottomrule
@@
 Discrete geometry studies the shapes and interactions between sets of discrete objects such as points, lines, circles, and polygons in 2-dimensional space.
-All of the discrete computioanl geomtery algorithms in compGeometeR take as an input a set of points $P$ in $n$-dimensional Eucliden space $\mathbb{R}^n$.
+All of the discrete computioanl geomtery algorithms in compGeometeR take as an input a set of points $P$ in $n$-dimensional Euclidean space $\mathbb{R}^n$.
 
@@
 Digital geometry is a branch of geometry that studies the geometric properties of a grid (or lattice) of points in Euclidean space, and usually involves the digitisation of discrete lines and regions \citep{rosenfeld-1989}.
+Digitisation occurs by representing Euclidean space $\mathbb{R}^n$ as a a rectangular orthogonal grid $\mathbb{G}^n$.
+The elements of $\mathbb{G}^2$ are called pixels, and the elements of $\mathbb{G}^3$ are called voxels \citep{klette-2004}.
+Each element in $\mathbb{G}^n$ has a grid coordinate for its centre, and a value that denotes if the element belongs to digitised geometric object, or when required which part of the digitised geometric object.
 
@@
 compGeometeR is very much a work in progress, and while there is sufficient functionality for it to be of wide use in the computational sciences, there is some functionality that has been identified as being particularly useful for future development.
+
+[Note inclusion of the Voronoi diagram for discrete geometry algorithms, and making more of the discrete algorithms available in digital form.]
 

### 2021-03-08 paper/main.tex
Updated paper

@@
 Digitisation occurs by representing Euclidean space $\mathbb{R}^n$ as a a rectangular orthogonal grid $\mathbb{G}^n$.
-The elements of $\mathbb{G}^2$ are called pixels, and the elements of $\mathbb{G}^3$ are called voxels \citep{klette-2004}.
-Each element in $\mathbb{G}^n$ has a grid coordinate for its centre, and a value that denotes if the element belongs to digitised geometric object, or when required which part of the digitised geometric object.
+The elements of $\mathbb{G}^2$ are called pixels, and the elements of $\mathbb{G}^3$ are called voxels, and each element in $\mathbb{G}^n$ has a grid coordinate for its centre \citep{klette-2004} and a value that denotes if the element belongs to a digitised geometric object, or when required, which part of the digitised geometric object .
 
@@
 compGeometeR is very much a work in progress, and while there is sufficient functionality for it to be of wide use in the computational sciences, there is some functionality that has been identified as being particularly useful for future development.
-
-[Note inclusion of the Voronoi diagram for discrete geometry algorithms, and making more of the discrete algorithms available in digital form.]
-
-\subsection{Geometric graph theory}
-% see: https://en.wikipedia.org/wiki/Geometric_graph_theory
-
-Geometric graph theory...
+The Voronoi diagram would be an obvious addition for the discrete geometry algorithms, as would making more of the discrete algorithms available in digital form.
 
@@
 
-\subsection{Spherical and ellipsoidal geometry}
-% see: https://en.wikipedia.org/wiki/Spherical_geometry
-% see: https://en.wikipedia.org/wiki/Elliptic_geometry
-
-All the algorithms in compGeometeR work in Cartesian space.
-
-\subsection{Fuzzy geometry}
-
 In some contexts it will be important to recognise that there is uncertainty in the position of points in space and the parameters defining a given algorithm.
@@
 Creating fuzzy versions of the compGeometeR algorithms could result in very useful functionality where quantifying uncertainty of any computational geometry is important.
-The existing digital geometric algorithms could be easily extended to fuzzy forms as all that is required is to
+The existing digital geometric algorithms could be easily extended to fuzzy forms as all that is required is to assign a membership value to each pixel \citep{klette-2004}.
+
+All the algorithms in compGeometeR work in Cartesian space, but there are logical extensions of these methodologies to spherical or ellipsoidal surfaces.
 

### 2021-03-09 paper/main.tex
Updated paper

@@
 
-\title{compGeometeR: an R package for computational geometry}
+\title{\texttt{compGeometeR}: an \texttt{R} package for computational geometry}
 
@@
 
-Computational geometry algorithms and data structures are widely applied across numerous scientific domains, and there a variety of R packages that apply computational geometry.
+Computational geometry algorithms and data structures are widely applied across numerous scientific domains, and there a variety of \texttt{R} packages that apply computational geometry.
 However, these packages often work in specific numbers of dimensions, have incompatible data structures, and include additional non-computational geometry functionality that can be domain specific.
-Our objective in developing the compGeometeR package is to produce a systematic implementation of the most commonly used computational geometry algorithms so that they can be easily integrated into domain specific scientific workflows.
-We briefly explain the algorithms available in compGeometeR, and identify priorities for future development.
+Our objective in developing the \texttt{compGeometeR} package is to produce a systematic implementation of the most commonly used computational geometry algorithms so that they can be easily integrated into domain specific scientific workflows.
+We briefly explain the algorithms available in \texttt{compGeometeR}, and identify priorities for future development.
 
 \begin{center}
-\textbf{Keywords}: R, alpha complex, alpha shape, convex hull, convex layers, Delaunay triangulation, digital, discrete.
+\textbf{Keywords}: alpha complex, alpha shape, convex hull, convex layers, Delaunay triangulation, digital, discrete.
 \end{center}
@@
 
-Our objective in developing the compGeometeR package is to produce a systematic implementation of some of the most commonly used computational geometry algorithms across various of branches of geometry so that they can be easily integrated into domain specific scientific workflows.
-The functionality of compGeometeR is entirely dependent on computational foundations provided by the Quickhull software \citep{barber-1996}.
+Our objective in developing the \texttt{compGeometeR} package is to produce a systematic implementation of some of the most commonly used computational geometry algorithms across various of branches of geometry so that they can be easily integrated into domain specific scientific workflows.
+The functionality of \texttt{compGeometeR} is entirely dependent on computational foundations provided by the \texttt{Qhull} C++ interface software \citep{barber-1996}.
 
@@
 
-There are a variety of computational geometry algorithms available in compGeometeR, some of which are replicated in other R packages.
-By documenting these differences (Table \ref{tab:r-options}) we hope to help potential users to assess if compGeometeR is the best option for their needs as there may be other options that are more suitable.
+There are a variety of computational geometry algorithms available in \texttt{compGeometeR}, some of which are replicated in other \texttt{R} packages.
+By documenting these differences (Table \ref{tab:r-options}) we hope to help potential users to assess if \texttt{compGeometeR} is the best option for their needs as there may be other options that are more suitable.
 
@@
 \centering
-\caption{Computational geometry software options in R.
-For each computational geometry function, the R software (and version) that can apply this algorithm is noted by reference to the number of dimensions in which the algorithm will work.}
+\caption{Computational geometry software options in \texttt{R}.
+For each computational geometry function, the \texttt{R} software (and version) that can apply this algorithm is noted by reference to the number of dimensions in which the algorithm will work.}
 \begin{tabular}{l c c c c c c c c}
@@
 
-   & \rotatebox{90}{compGeometeR (1.0.0)}
-   & \rotatebox{90}{R (3.5.3)}
-   & \rotatebox{90}{alphahull (2.2)}
-   & \rotatebox{90}{alphashape3d (1.3.1)}
-   & \rotatebox{90}{deldir (0.1-25)}
-   & \rotatebox{90}{geometry (0.4.5)}
-   & \rotatebox{90}{spatstat (1.63-3)}
-   & \rotatebox{90}{tripack (1.3-9.1)} \\
+   & \rotatebox{90}{\texttt{compGeometeR} (1.0.0)}
+   & \rotatebox{90}{\texttt{R} (3.5.3)}
+   & \rotatebox{90}{\texttt{alphahull} (2.2)}
+   & \rotatebox{90}{\texttt{alphashape3d} (1.3.1)}
+   & \rotatebox{90}{\texttt{deldir} (0.1-25)}
+   & \rotatebox{90}{\texttt{geometry} (0.4.5)}
+   & \rotatebox{90}{\texttt{spatstat} (1.63-3)}
+   & \rotatebox{90}{\texttt{tripack} (1.3-9.1)} \\
 
@@
 Discrete geometry studies the shapes and interactions between sets of discrete objects such as points, lines, circles, and polygons in 2-dimensional space.
-All of the discrete computioanl geomtery algorithms in compGeometeR take as an input a set of points $P$ in $n$-dimensional Euclidean space $\mathbb{R}^n$.
+All of the discrete computational geometry algorithms in \texttt{compGeometeR} take as an input a set of points $P$ in $n$-dimensional Euclidean space $\mathbb{R}^n$.
 
@@
 \includegraphics{figures/figure-1/figure-1.png}
-\caption{Two-dimensional examples of discrete geometry algorithms currently available in compGeometeR. (a) Convex hull, (b) convex layers, (c) Delaunay triangulation, and (d) alpha complex.}
+\caption{Two-dimensional examples of discrete geometry algorithms currently available in \texttt{compGeometeR}. (a) Convex hull, (b) convex layers, (c) Delaunay triangulation, and (d) alpha complex.}
 \label{fig:euclidean-algorithms}
@@
 
-compGeometeR is very much a work in progress, and while there is sufficient functionality for it to be of wide use in the computational sciences, there is some functionality that has been identified as being particularly useful for future development.
-The Voronoi diagram would be an obvious addition for the discrete geometry algorithms, as would making more of the discrete algorithms available in digital form.
+\texttt{compGeometeR} is very much a work in progress, and while there is sufficient functionality for it to be of wide use in the computational sciences, there is some functionality that has been identified as being particularly useful for future development.
+
+The Voronoi diagram \citep{voronoi-1908, okabe-2000} partitions $n$-dimensional space into a set of polytopes for which each polytope delineates the region of $n$-dimensional space that is closest to each point in $P$.
+The Voronoi diagram would be an obvious addition for the discrete geometry algorithms given its wide implementation in other R geometry packages (Table \ref{tab:r-options}).
+
+As well as expanding the number of discrete geometry algorithms, and implementing more discrete algorithms in digital form, there are other kinds of computational geometry algorithms that we think would be of particular value.
 
@@
 In such situations it may be helpful to adopt fuzzy geometry \citep{rosenfeld-1998} view in which, foe example, memberships of points in space are not crisp being either 0 or 1, but rather are fuzzy on a scale from 0 to 1.
-Creating fuzzy versions of the compGeometeR algorithms could result in very useful functionality where quantifying uncertainty of any computational geometry is important.
+Creating fuzzy versions of the \texttt{compGeometeR} algorithms could result in very useful functionality where quantifying uncertainty of any computational geometry is important.
 The existing digital geometric algorithms could be easily extended to fuzzy forms as all that is required is to assign a membership value to each pixel \citep{klette-2004}.
 
-All the algorithms in compGeometeR work in Cartesian space, but there are logical extensions of these methodologies to spherical or ellipsoidal surfaces.
+All the algorithms in \texttt{compGeometeR} work in Euclidean space, but computational geometry algorithms can also be usefully applied in spherical or hyperbolic space \citep{gowers-2003}.
+For example, the Fortran \texttt{STRIPACK} software \citep{renka-1997} could be wrapped by R to provide functionality to compute Delaunay triangulations and Voronoi diagrams on the surface of a sphere.
 
@@
 
-compGeometeR is open source software made available under the GPL license. Installation
-instructions can be found at the GitHub repository \url{https://github.com/manaakiwhenua/compGeometeR}.
+\texttt{compGeometeR} is open source software made available under a General Public License license.
+Installation instructions can be found at the GitHub repository \url{https://github.com/manaakiwhenua/compGeometeR}.
 

### 2021-03-09 R/grid-coordinates.R
Updated grid_coordinates to include dimension coordinates

@@ -34,11 +34,15 @@ grid_coordinates <- function(mins, maxs, spacings) {
   }
   
   # Create list of coordinate locations for each dimension
-  dimCoords = list()
-  for (n in seq(1, length(mins))) {
-    dimCoords[[n]] <- seq(mins[n], maxs[n], spacings[n])
+  dims = length(mins)
+  dimension_coords <- list()
+  for (n in seq(dims)) {
+    dimension_coords[[n]] <- seq(mins[n], maxs[n], spacings[n])
   }
   # Create all combinations of coordinates across all dimension
-  return(expand.grid(dimCoords))
+  grid_coords <- expand.grid(dimension_coords, KEEP.OUT.ATTRS = FALSE)
+  colnames(grid_coords) <- seq(dims)
+  
+  return(list(grid_coords, dimension_coords))
   
 }

### 2021-03-09 R/in-convex-hull.R
Updated in_convex_hull to return integer rather than boolean values

@@ -10,9 +10,9 @@
 #'   represent \eqn{n} points and the \eqn{d} columns the coordinates in 
 #'   \eqn{d}-dimensional space.
 #' 
-#' @return A \eqn{n} length vector containing \code{TRUE} if test point \eqn{n} 
-#' lies within the hull and \code{FALSE} if it lies outside the hull.  If any of 
-#' the test point coordinates contain NA then the output is also NA.
+#' @return A \eqn{n} length vector containing 1 if test point \eqn{n} 
+#' lies within the hull and 0 if it lies outside the hull.  If any of 
+#' the test point coordinates contain NA then the output is 0.
 #' 
 #' @seealso \code{\link{convex_hull}}
 #' 
@@ -87,6 +87,6 @@
     # Replace any test point coordinates containing NAs with NA
     in_hull[!complete.cases(test_points)] = NA
     
-    return(in_hull)
+    return(as.integer(in_hull))
     
   }

### 2021-03-09 R/grid-coordinates.R
Updated function descriptions

@@ -10,11 +10,14 @@
 #' @param spacings Vector of length \code{n} listing the spacings between points 
 #' in all dimensions.
 #'
-#' @return A matrix with \code{n} columns - and potentially lots of rows!
+#' @return A list of two objects.  The first object is a dataframe with \code{n} 
+#' columns and a row for each coordinate - so potentially lots of rows!  The 
+#' second object is another list of length \code{n} that contains the 
+#' coordinates along each dimension.
 #'
 #' @examples
 #' # Point space grid coordinates usage
-#' xy = grid_coordinates(mins=c(0,0), maxs=c(10,15), spacings=c(1,1))
+#' grid = grid_coordinates(mins=c(0,0), maxs=c(10,15), spacings=c(1,1))
 #' 
 #' @export
 grid_coordinates <- function(mins, maxs, spacings) {

### 2021-03-09 R/convex-hull.R
Updated convex_hull fucntion

@@ -22,6 +22,10 @@
 #'   convex hull.
 #' }
 #' 
+#' In the \eqn{2}-dimensional case the convex hull indices and vertices are 
+#' returned in a circular order to ease plotting, but in other dimensions there 
+#' is no specific order.
+#' 
 #' @seealso \code{\link{convex_layer}}
 #' 
 #' @references Barber CB, Dobkin DP, Huhdanpaa H (1996) The Quickhull algorithm 
@@ -82,6 +86,16 @@
   	convex$hull_simplices <- as.matrix(simplices)
   	convex$hull_indices <- unique(c(as.integer(ch$convex_hull + 1)))
   	convex$hull_vertices <- points[convex$hull_indices,]
+  	
+  	# If the convex hull is 2-dimensional sort the vertices in a circular order
+  	if (ncol(points) == 2) {
+  	  midpoint <- colMeans(convex$hull_vertices)
+      angles <- atan2(convex$hull_vertices[,1] - midpoint[1], convex$hull_vertices[,2] - midpoint[2])
+      angles[angles < 0] <- angles[angles < 0] + 2 * pi
+      ch_vertices_order <- sort(angles, index.return=TRUE)$ix
+      convex$hull_indices <- convex$hull_indices[ch_vertices_order,]
+  	  convex$hull_vertices <- points[convex$hull_indices,]
+  	}
   
   	return(convex)
   }

### 2021-03-09 R/convex-hull.R
Updated convex_hull fucntion

@@ -93,7 +93,8 @@
       angles <- atan2(convex$hull_vertices[,1] - midpoint[1], convex$hull_vertices[,2] - midpoint[2])
       angles[angles < 0] <- angles[angles < 0] + 2 * pi
       ch_vertices_order <- sort(angles, index.return=TRUE)$ix
-      convex$hull_indices <- convex$hull_indices[ch_vertices_order,]
+      
+      convex$hull_indices <- convex$hull_indices[ch_vertices_order]
   	  convex$hull_vertices <- points[convex$hull_indices,]
   	}
   

### 2021-03-09 R/convex-hull.R
Updated convex_hull function decription and example

@@ -40,9 +40,7 @@
 #' # Create convex hull and plot
 #' ch <- convex_hull(points=p)
 #' plot(p, pch = as.character(seq(nrow(p))))
-#' for (s in seq(nrow(ch$hull_simplices))) {
-#'   lines(ch$input_points[ch$hull_simplices[s, ], ], col = "red")
-#' }
+#' polygon(ch$hull_vertices, border="red")
 #' 
 #' @export
   convex_hull <- function(points=NULL) {

### 2021-03-09 R/find-simplex.R
Updated find_simplex to return 0 rather than NA

@@ -11,9 +11,9 @@
 #'   \eqn{d}-dimensional space. 
 #' 
 #' @return A \eqn{n} length vector containing the index of the simplex the test 
-#' point is within, or a value of NA if a test point is not within any of the 
+#' point is within, or a value of 0 if a test point is not within any of the 
 #' simplices.  If any of the test point coordinates contain NA then the output 
-#' is also NA.
+#' is also 0.
 #' 
 #' @examples 
 #' # Define points and create an alpha complex
@@ -63,7 +63,7 @@ find_simplex <- function(simplices, test_points) {
   inHull_test <- as.matrix(test_points[inHull_test_point_indices,])
 
   # Create an empty object to hold which simplex the test points are in
-  test_points_simplex <- rep(NA, nrow(test_points))
+  test_points_simplex <- rep(0, nrow(test_points))
   # For each simplex
   for(simplex in c(1:nrow(simplices$simplices))) {
     

### 2021-03-10 R/digital-convex-hull.R
Added digital_convex_hull function

@@ -1,0 +1,59 @@
+#' @title Digital convex hull
+#' 
+#' @description  This function calculates the digital 
+#' \href{https://en.wikipedia.org/wiki/Convex_hull}{convex hull} around a set of
+#' \eqn{n} points in \eqn{d}-dimensional space based upon a grid of 
+#' \eqn{d}-dimensional coordinates.
+#'
+#' @param points a \eqn{n}-by-\eqn{d} dataframe or matrix. The rows
+#'   represent \eqn{n} points and the \eqn{d} columns the coordinates in 
+#'   \eqn{d}-dimensional space.
+#' @param mins Vector of length \code{d} listing the grid coordinate minimum for 
+#' each dimension.
+#' @param maxs Vector of length \code{d} listing the grid coordinate maximum for 
+#' each dimension.
+#' @param spacings Vector of length \code{d} listing the grid coordinate spacing 
+#' for each dimension.
+#' 
+#' @return A list of three objects:
+#' 
+#' \itemize{
+#'   \item A \eqn{d}-dimensional array containing 1 if a grid coordinate lies 
+#'   within the hull and 0 if it lies outside the hull (if any of the test point 
+#'   coordinates contain NA then the output is 0).
+#'   \item A dataframe with \code{d} columns and a row for each grid coordinate 
+#'   - so potentially lots of rows!
+#'   \item A list of length \code{d} that contains the grid coordinates along 
+#'   each dimension.
+#' }
+#'
+#' @examples
+#' # Define points
+#' x <- c(30, 70, 20, 50, 40, 70)
+#' y <- c(35, 80, 70, 50, 60, 20)
+#' p <- data.frame(x, y)
+#' # Create digital convex hull and plot
+#' d_ch <- digital_convex_hull(points = p, mins=c(15,15), maxs=c(85,85), spacings=c(0.5,0.5))
+#' image(x=d_ch[[3]][[1]], y=d_ch[[3]][[2]], z=d_ch[[1]], xlab="x", ylab="y")
+#' points(p, pch = as.character(seq(nrow(p))))
+#' 
+#' @export
+digital_convex_hull <- function(points = p, mins, maxs, spacings) {
+
+  # Create the discrete convex hull
+  ch <- convex_hull(points=p)
+  # Generate a grid of coordinates
+  grid <- grid_coordinates(mins, maxs, spacings)
+  # Check which grid coordinates are in the convex hull
[…]

### 2021-03-10 R/grid-coordinates.R

@@ -1,19 +1,23 @@
 #' @title Grid Coordinates
 #' 
-#' @description This function creates a \code{n}-dimensional grid of coordinates 
+#' @description This function creates a \code{d}-dimensional grid of coordinates 
 #' systematically spaced throughout the specified spatial extent.
 #'
-#' @param mins Vector of length \code{n} listing the point space minimum for each
-#' dimension.
-#' @param maxs Vector of length \code{n} listing the point space maximum for each
-#' dimension.
-#' @param spacings Vector of length \code{n} listing the spacings between points 
-#' in all dimensions.
+#' @param mins Vector of length \code{d} listing the grid coordinate minimum for 
+#' each dimension.
+#' @param maxs Vector of length \code{d} listing the grid coordinate maximum for 
+#' each dimension.
+#' @param spacings Vector of length \code{d} listing the grid coordinate spacing 
+#' for each dimension.
 #'
-#' @return A list of two objects.  The first object is a dataframe with \code{n} 
-#' columns and a row for each coordinate - so potentially lots of rows!  The 
-#' second object is another list of length \code{n} that contains the 
-#' coordinates along each dimension.
+#' @return A list of two objects:
+#' 
+#' \itemize{
+#'   \item A dataframe with \code{d} columns and a row for each grid coordinate 
+#'   - so potentially lots of rows!
+#'   \item A list of length \code{d} that contains the grid coordinates along 
+#'   each dimension.
+#' }
 #'
 #' @examples
 #' # Point space grid coordinates usage

### 2021-03-10 R/digital-alpha-complex.R
Added digital_alpha_complex function

@@ -1,13 +1,17 @@
 #' @title Digital convex hull
 #' 
 #' @description  This function calculates the digital 
-#' \href{https://en.wikipedia.org/wiki/Convex_hull}{convex hull} around a set of
-#' \eqn{n} points in \eqn{d}-dimensional space based upon a grid of 
+#' \href{https://en.wikipedia.org/wiki/Alpha_shape#Alpha_complex}{alpha complex}
+#' of a set of \eqn{n} points in \eqn{d}-dimensional space based upon a grid of 
 #' \eqn{d}-dimensional coordinates.
 #'
 #' @param points a \eqn{n}-by-\eqn{d} dataframe or matrix. The rows
 #'   represent \eqn{n} points and the \eqn{d} columns the coordinates in 
 #'   \eqn{d}-dimensional space.
+#' @param alpha a real number between zero and infinity that defines the maximum 
+#'   circumradii for a simplex to be included in the alpha complex.  If 
+#'   unspecified \code{alpha} defaults to infinity and the alpha complex is 
+#'   equivalent to a Delaunay triangulation.
 #' @param mins Vector of length \code{d} listing the grid coordinate minimum for 
 #' each dimension.
 #' @param maxs Vector of length \code{d} listing the grid coordinate maximum for 
@@ -18,9 +22,10 @@
 #' @return A list of three objects:
 #' 
 #' \itemize{
-#'   \item A \eqn{d}-dimensional array containing 1 if a grid coordinate lies 
-#'   within the hull and 0 if it lies outside the hull (if any of the test point 
-#'   coordinates contain NA then the output is 0).
+#'   \item A \eqn{d}-dimensional array containing an integer index of the alpha 
+#'   complex \eqn{s} \href{https://en.wikipedia.org/wiki/Simplex}{simplex} that 
+#'   each grid coordinate lies within, or 0 if it lies outside the alpha complex 
+#'   (if any of the test point coordinates contain NA then the output is 0).
 #'   \item A dataframe with \code{d} columns and a row for each grid coordinate 
 #'   - so potentially lots of rows!
 #'   \item A list of length \code{d} that contains the grid coordinates along 
@@ -38,7 +43,7 @@
 #' points(p, pch = as.character(seq(nrow(p))))
 #' 
 #' @export
-digital_convex_hull <- function(points=NULL, mins, maxs, spacings) {
+digital_convex_hull <- function(points, mins, maxs, spacings) {
 
   # Create the discrete convex hull
   ch <- convex_hull(points=p)

### 2021-03-10 R/find-simplex.R

@@ -90,11 +90,12 @@ find_simplex <- function(simplices, test_points) {
       barycentric_coords <- (to_test - matrix(simplex_coordinates[dim+1,], n, dim, byrow=TRUE)) %*% solve(X1)
       barycentric_coords <- cbind(barycentric_coords, 1 - apply(barycentric_coords, 1, sum))
       
-      # Those test points for which are coords are positive are in the simplex
+      # Those test points for which are coords are non-negative are in the simplex
       barycentric_sign <- sign(barycentric_coords)
+      barycentric_sign[barycentric_sign == 0] = 1
       barycentric_sign_sum <- rowSums(barycentric_sign)
-      in_simplex = which(barycentric_sign_sum == d + 1)
-      test_points_simplex[as.numeric(row.names(to_test[in_simplex, , drop=FALSE]))] = simplex
+      in_simplex <- which(barycentric_sign_sum == d + 1)
+      test_points_simplex[as.numeric(row.names(to_test[in_simplex, , drop=FALSE]))] <- simplex
     }
   }
   

### 2021-03-10 R/digital-alpha-complex.R
Updated digital_alpha_complex function

@@ -1,4 +1,4 @@
-#' @title Digital convex hull
+#' @title Digital alpha complex
 #' 
 #' @description  This function calculates the digital 
 #' \href{https://en.wikipedia.org/wiki/Alpha_shape#Alpha_complex}{alpha complex}
@@ -37,28 +37,32 @@
 #' x <- c(30, 70, 20, 50, 40, 70)
 #' y <- c(35, 80, 70, 50, 60, 20)
 #' p <- data.frame(x, y)
-#' # Create digital convex hull and plot
-#' d_ch <- digital_convex_hull(points = p, mins=c(15,15), maxs=c(85,85), spacings=c(0.5,0.5))
-#' image(x=d_ch[[3]][[1]], y=d_ch[[3]][[2]], z=d_ch[[1]], xlab="x", ylab="y")
+#' # Create digital alpha complex and plot
+#' d_ac <- digital_alpha_complex(points = p, alpha = 20, mins=c(15,15), maxs=c(85,85), spacings=c(0.5,0.5))
+#' cols = c("lightgrey", "orange", "purple", "lightseagreen")
+#' image(x=d_ac[[3]][[1]], y=d_ac[[3]][[2]], z=d_ac[[1]], xlab="x", ylab="y", col = cols)
+#' points(p, pch = as.character(seq(nrow(p))))
+#' legend("bottomleft", pch=15, col=cols, legend=sort(unique(c(d_ac[[1]]))), title="Simplex")
 #' points(p, pch = as.character(seq(nrow(p))))
 #' 
 #' @export
-digital_convex_hull <- function(points, mins, maxs, spacings) {
+digital_alpha_complex <- function(points=NULL, alpha=Inf, mins, maxs, spacings) {
 
-  # Create the discrete convex hull
-  ch <- convex_hull(points=p)
+  # Create the discrete alpha complex
+  ac <- alpha_complex(points = p, alpha = alpha)
   # Generate a grid of coordinates
   grid <- grid_coordinates(mins, maxs, spacings)
-  # Check which grid coordinates are in the convex hull
-  m <- in_convex_hull(ch, grid[[1]])
+  # Check which simplex the grid coordinates are in
+  m <- find_simplex(ac, grid[[1]])
   # Get the grid length of each dimension
   dim_n <- c()
   for (dim in grid[[2]]) {
     dim_n <- c(dim_n, length(dim))
   }
   # Create an array of the results
[…]

### 2021-03-10 R/digital-alpha-shape.R
Added digital_alpha_shape function

@@ -1,0 +1,65 @@
+#' @title Digital alpha shape
+#' 
+#' @description  This function calculates the digital 
+#' \href{https://en.wikipedia.org/wiki/Alpha_shape}{alpha shape}
+#' of a set of \eqn{n} points in \eqn{d}-dimensional space based upon a grid of 
+#' \eqn{d}-dimensional coordinates.
+#'
+#' @param points a \eqn{n}-by-\eqn{d} dataframe or matrix. The rows
+#'   represent \eqn{n} points and the \eqn{d} columns the coordinates in 
+#'   \eqn{d}-dimensional space.
+#' @param alpha a real number between zero and infinity that defines the maximum 
+#'   circumradii for a simplex to be included in the alpha shape  If 
+#'   unspecified \code{alpha} defaults to infinity and the alpha shape is 
+#'   equivalent to a Delaunay triangulation.
+#' @param mins Vector of length \code{d} listing the grid coordinate minimum for 
+#' each dimension.
+#' @param maxs Vector of length \code{d} listing the grid coordinate maximum for 
+#' each dimension.
+#' @param spacings Vector of length \code{d} listing the grid coordinate spacing 
+#' for each dimension.
+#' 
+#' @return A list of three objects:
+#' 
+#' \itemize{
+#'   \item A \eqn{d}-dimensional array containing 1 if a grid coordinate lies 
+#'   inside the alpha shape and 0 if it lies outside (if any of the test point 
+#'   coordinates contain NA then the output is 0).
+#'   \item A dataframe with \code{d} columns and a row for each grid coordinate 
+#'   - so potentially lots of rows!
+#'   \item A list of length \code{d} that contains the grid coordinates along 
+#'   each dimension.
+#' }
+#'
+#' @examples
+#' # Define points
+#' x <- c(30, 70, 20, 50, 40, 70)
+#' y <- c(35, 80, 70, 50, 60, 20)
+#' p <- data.frame(x, y)
+#' # Create digital alpha complex and plot
+#' d_as <- digital_alpha_shape(points = p, alpha = 20, mins=c(15,15), maxs=c(85,85), spacings=c(0.5,0.5))
+#' image(x=d_as[[3]][[1]], y=d_as[[3]][[2]], z=d_as[[1]], xlab="x", ylab="y")
+#' points(p, pch = as.character(seq(nrow(p))))
+#' 
+#' @export
[…]

### 2021-03-10 paper/main.tex
Updated paper

@@
 
-A convex hull defines the smallest subset of $\mathbb{R}^n$ that contains $P$ and for which the subset is convex so any two points within the convex hull can be connected by a straight line also contained by the convex hull (Figure \ref{fig:euclidean-algorithms}a).
+A convex hull \citep{barber-1996} defines the smallest subset of $\mathbb{R}^n$ that contains $P$ and for which the subset is convex so any two points within the convex hull can be connected by a straight line also contained by the convex hull (Figure \ref{fig:discrete-algorithms}a).
 
@@
 Convex layers are a nested sequence of convex hulls produced by repeating the process of constructing a convex hull for $P$ and then removing the points forming the vertices of the convex hull from $P$ before producing the next convex layer.
-The first convex layer is equivalent to the convex hull, with each successive convex layer representing ever smaller region of space (Figure \ref{fig:euclidean-algorithms}b).
+The first convex layer is equivalent to the convex hull, with each successive convex layer representing ever smaller region of space (Figure \ref{fig:discrete-algorithms}b).
 
-The Delaunay triangulation \citep{delaunay-1934} produces a set of simplices (triangles in 2-dimensions or tetrahedrons in 3-dimensions) for which no point in $P$ is inside the the circumhypersphere (a circle in 2-dimensions or a sphere in 3-dimensions) of each simplex (Figure \ref{fig:euclidean-algorithms}c).
+The Delaunay triangulation \citep{delaunay-1934} produces a set of simplices (triangles in 2-dimensions or tetrahedrons in 3-dimensions) for which no point in $P$ is inside the the circumhypersphere (a circle in 2-dimensions or a sphere in 3-dimensions) of each simplex (Figure \ref{fig:discrete-algorithms}c).
 
-The alpha complex \citep{edelsbrunner-1994} is a subset of the Delaunay triangulation that contains only those simplices for whose circumhypersphere radius is smaller than a specified $\alpha$ parameter value ranging $0 < \alpha < \infty$ (Figure \ref{fig:euclidean-algorithms}d).
+The alpha complex \citep{edelsbrunner-1994} is a subset of the Delaunay triangulation that contains only those simplices for whose circumhypersphere radius is smaller than a specified $\alpha$ parameter value ranging $0 < \alpha < \infty$ (Figure \ref{fig:discrete-algorithms}d).
+The related alpha shape would be a polytope that combines all of the simplices in an alpha complex \citep{edelsbrunner-1994}.
 
@@
 \caption{Two-dimensional examples of discrete geometry algorithms currently available in \texttt{compGeometeR}. (a) Convex hull, (b) convex layers, (c) Delaunay triangulation, and (d) alpha complex.}
-\label{fig:euclidean-algorithms}
+\label{fig:discrete-algorithms}
 \end{figure}
@@
 Digitisation occurs by representing Euclidean space $\mathbb{R}^n$ as a a rectangular orthogonal grid $\mathbb{G}^n$.
-The elements of $\mathbb{G}^2$ are called pixels, and the elements of $\mathbb{G}^3$ are called voxels, and each element in $\mathbb{G}^n$ has a grid coordinate for its centre \citep{klette-2004} and a value that denotes if the element belongs to a digitised geometric object, or when required, which part of the digitised geometric object .
+The elements of $\mathbb{G}^2$ are called pixels, and the elements of $\mathbb{G}^3$ are called voxels, and each element in $\mathbb{G}^n$ has a grid coordinate for its centre \citep{klette-2004} and a value that denotes if the element belongs to a digitised geometric object, or when required, which part of the digitised geometric object.
+\texttt{compGeometeR} has implemented digital versions of the convex hull (Figure \ref{fig:digital-algorithms}a), alpha complex (Figure \ref{fig:digital-algorithms}b), and alpha shape (Figure \ref{fig:digital-algorithms}c).
+
+\begin{figure}[ht]
+\centering
+\includegraphics{figures/figure-2/figure-2.png}
+\caption{Two-dimensional examples of digital geometry algorithms currently available in \texttt{compGeometeR}. (a) Convex hull, (b) alpha complex, and (c) alpha shape.}
+\label{fig:digital-algorithms}
+\end{figure}
 

### 2021-03-10 paper/figures/figure-1/figure-1.R

@@ -14,25 +14,13 @@ p = cbind(x, y)
 
 #-------------------------------------------------------------------------------
 
-# Function to sort vertices into a circular order
-circular_vertex_sort <- function(vertices) {
-  midpoint = colMeans(vertices)
-  angles = atan2(vertices[,1] - midpoint[1], vertices[,2] - midpoint[2])
-  angles[angles < 0] = angles[angles < 0] + 2 * pi
-  ch_vertices_order = sort(angles, index.return=TRUE)$ix
-  vertices[ch_vertices_order,]
-}
-
-#-------------------------------------------------------------------------------
-
 png("figure-1.png", width = 10, height = 10, res=300, units = "cm", pointsize = 10)
 par(mfrow=c(2,2), oma=c(0,0,0,0), mar=c(0.5,0.5,1.5,0.5))
 
 # Convex hull
 plot(x, y, yaxt="n", xaxt="n", xlab="", ylab = "", pch=16, cex=0.75)
 ch = convex_hull(p)
-circ_vertices = circular_vertex_sort(ch$hull_vertices)
-polygon(circ_vertices, col="lightgrey", border="darkgrey")
+polygon(ch$hull_vertices, col="orange", border="firebrick")
 points(p, pch=16, cex=0.75)
 mtext("(a)", side = 3, line=0.25)
 
@@ -40,8 +28,7 @@ mtext("(a)", side = 3, line=0.25)
 plot(x, y, yaxt="n", xaxt="n", xlab="", ylab = "", pch=16, cex=0.75)
 for (i in seq(4)) {
   cl = convex_layer(p, layer = i)
-  circ_vertices = circular_vertex_sort(cl$hull_vertices)
-  polygon(circ_vertices, col="lightgrey", border="darkgrey")
+  polygon(cl$hull_vertices, col="orange", border="firebrick")
 }
 points(p, pch=16, cex=0.75)
 mtext("(b)", side = 3, line=0.25)
@@ -50,7 +37,7 @@ mtext("(b)", side = 3, line=0.25)
 plot(x, y, yaxt="n", xaxt="n", xlab="", ylab = "", pch=16, cex=0.75)
 dt = delaunay(p)
 for (s in seq(nrow(dt$simplices))) {
-  polygon(dt$input_points[dt$simplices[s,],], col="lightgrey", border="darkgrey")
[…]

### 2021-03-10 paper/figures/figure-2/figure-2.R

@@ -1,0 +1,47 @@
+#-------------------------------------------------------------------------------
+
+# R version 3.5.3 (2019-03-11) -- "Great Truth"
+# Platform: x86_64-w64-mingw32/x64 (64-bit)
+
+library(compGeometeR) # version 1.0
+
+#-------------------------------------------------------------------------------
+
+set.seed(2) # to reproduce figure exactly
+x = rgamma(n = 20, shape = 3, scale = 2)
+y = rnorm(n = 20, mean = 10, sd = 2)
+p = cbind(x, y)
+
+#-------------------------------------------------------------------------------
+
+png("figure-2.png", width = 15, height = 5, res=300, units = "cm", pointsize = 10)
+par(mfrow=c(1,3), oma=c(0,0,0,0), mar=c(0.5,0.5,1.5,0.5))
+cols = c("white", "slateblue", "firebrick", "lightseagreen", "purple", "darkorchid", "dodgerblue",
+         "goldenrod", "grey", "deeppink", "turquoise", "yellowgreen", "yellow", "limegreen",
+         "red2", "navajowhite2", "orange", "darkgrey")
+
+# Digital convex hull
+d_ch = digital_convex_hull(p, mins=c(0,5), maxs=c(15,15), spacings = c(0.05,0.05))
+image(x=d_ch[[3]][[1]], y=d_ch[[3]][[2]], z=d_ch[[1]], 
+      yaxt="n", xaxt="n", xlab="", ylab = "", col=c("white", "orange"))
+points(p, pch=16, cex=0.75)
+mtext("(a)", side = 3, line=0.25)
+
+# Digital alpha complex
+d_ac = digital_alpha_complex(p, alpha = 2, mins=c(0,5), maxs=c(15,15), spacings = c(0.05,0.05))
+image(x=d_ac[[3]][[1]], y=d_ac[[3]][[2]], z=d_ac[[1]]+1, 
+      yaxt="n", xaxt="n", xlab="", ylab = "", col=cols)
+points(p, pch=16, cex=0.75)
+mtext("(b)", side = 3, line=0.25)
+
+# Digital alpha shape
+d_as = digital_alpha_shape(p, alpha = 2, mins=c(0,5), maxs=c(15,15), spacings = c(0.05,0.05))
+image(x=d_as[[3]][[1]], y=d_as[[3]][[2]], z=d_as[[1]], 
+      yaxt="n", xaxt="n", xlab="", ylab = "", col=c("white", "orange"))
[…]

### 2021-03-10 paper/main.tex
Updated paper

@@
 \usepackage{booktabs}
-\usepackage{hyperref} % to insert a web link
+\usepackage[hidelinks]{hyperref} % to insert a web link
 \usepackage{amsfonts} % to get blackboard font
+
+\usepackage{listings} % to get code blocks
+\usepackage{xcolor}
+\definecolor{codegreen}{rgb}{0,0.6,0}
+\definecolor{codegray}{rgb}{0.5,0.5,0.5}
+\definecolor{codepurple}{rgb}{0.58,0,0.82}
+\definecolor{backcolour}{rgb}{0.9,0.9,0.9}
+\lstdefinestyle{mystyle}{
+    backgroundcolor=\color{backcolour},   
+    commentstyle=\color{codegreen},
+    keywordstyle=\color{magenta},
+    numberstyle=\tiny\color{codegray},
+    stringstyle=\color{codepurple},
+    basicstyle=\ttfamily\footnotesize,
+    breakatwhitespace=false,         
+    breaklines=true,                 
+    captionpos=b,                    
+    keepspaces=true,                 
+    numbers=left,                    
+    numbersep=5pt,                  
+    showspaces=false,                
+    showstringspaces=false,
+    showtabs=false,                  
+    tabsize=2
+}
+\lstset{style=mystyle}
 
@@
 Computational geometry seeks to develop algorithms and data structures for geometric objects, and the results are widely applied across numerous scientific domains \citep{de-berg-2008}.
-Given the importance of computational geometry it is no surprise that within the R computing ecosystem there are a variety of options for applying various computational geometry algorithms in a variety of dimensions.
+Given the importance of computational geometry it is no surprise that within the \texttt{R} computing ecosystem there are a variety of options for applying various computational geometry algorithms.
 However, these packages often work in specific numbers of dimensions, have incompatible data structures, and include additional non-computational geometry functionality that can be domain specific.
@@
 Our objective in developing the \texttt{compGeometeR} package is to produce a systematic implementation of some of the most commonly used computational geometry algorithms across various of branches of geometry so that they can be easily integrated into domain specific scientific workflows.
-The functionality of \texttt{compGeometeR} is entirely dependent on computational foundations provided by the \texttt{Qhull} C++ interface software \citep{barber-1996}.
 
@@
 By documenting these differences (Table \ref{tab:r-options}) we hope to help potential users to assess if \texttt{compGeometeR} is the best option for their needs as there may be other options that are more suitable.
+
+The functions of \texttt{compGeometeR} interconnect and build upon one another, but it is important to recognise that virtually all of the \texttt{compGeometeR} functions are  dependent on computational geometry foundations provided by the \texttt{Qhull} C++ interface software \citep{barber-1996}.
 
@@
 
+\begin{figure}[ht]
+\centering
+\includegraphics[width=15cm]{figures/software-structure/software-structure.png}
+\caption{Connections and dependencies of the \texttt{compGeometeR} functions.}
+\label{fig:software-structure}
+\end{figure}
+
 \subsection{Discrete geometry}
@@
 A convex hull \citep{barber-1996} defines the smallest subset of $\mathbb{R}^n$ that contains $P$ and for which the subset is convex so any two points within the convex hull can be connected by a straight line also contained by the convex hull (Figure \ref{fig:discrete-algorithms}a).
+As an example of the simplicity of \texttt{compGeometeR} a convex hull can be generated and visualised with minimal code (Listing \ref{code:discrete-convex-hull-code}).
+
+\begin{lstlisting}[language=R, caption=Example \texttt{R} code to create a discrete convex hull with \texttt{compGeometeR}, label={code:discrete-convex-hull-code}]
+library(compGeometeR)
+# Generate point data
+set.seed(2) # to reproduce figure exactly
+x = rgamma(n = 20, shape = 3, scale = 2)
+y = rnorm(n = 20, mean = 10, sd = 2)
+p = cbind(x, y)
+# Create convex hull
+ch = convex_hull(p)
+# Plot point data and convex hull
+plot(x, y, yaxt="n", xaxt="n", xlab="", ylab = "", pch=16, cex=0.75)
+polygon(ch$hull_vertices, col="orange", border="firebrick")
+points(p, pch=16, cex=0.75)
+\end{lstlisting}
 
@@
 \centering
-\includegraphics{figures/figure-1/figure-1.png}
+\includegraphics{figures/discrete-algorithms/discrete-algorithms.png}
 \caption{Two-dimensional examples of discrete geometry algorithms currently available in \texttt{compGeometeR}. (a) Convex hull, (b) convex layers, (c) Delaunay triangulation, and (d) alpha complex.}
@@
 The elements of $\mathbb{G}^2$ are called pixels, and the elements of $\mathbb{G}^3$ are called voxels, and each element in $\mathbb{G}^n$ has a grid coordinate for its centre \citep{klette-2004} and a value that denotes if the element belongs to a digitised geometric object, or when required, which part of the digitised geometric object.
-\texttt{compGeometeR} has implemented digital versions of the convex hull (Figure \ref{fig:digital-algorithms}a), alpha complex (Figure \ref{fig:digital-algorithms}b), and alpha shape (Figure \ref{fig:digital-algorithms}c).
+\texttt{compGeometeR} has implemented digital versions of the convex hull (Figure \ref{fig:digital-algorithms}a), alpha complex (Figure \ref{fig:digital-algorithms}b), and alpha shape (Figure \ref{fig:digital-algorithms}c) that are unique functions amongst \texttt{R} software (Table \ref{tab:r-options}).
+
+The code required for digital versions of the discrete algorithms is no more complicated, and simply requires addtional parameters to define the extent and resolution of $\mathbb{G}^n$.
+For example, the code required for a digital convex hull (Listing \ref{code:digital-convex-hull-code}) does not differ much from that required for a discrete convex hull (Listing \ref{code:discrete-convex-hull-code}).
+
+\begin{lstlisting}[language=R, caption=Example \texttt{R} code to create a digital convex hull with \texttt{compGeometeR}, label={code:digital-convex-hull-code}]
+library(compGeometeR)
+# Generate point data
+set.seed(2) # to reproduce figure exactly
+x = rgamma(n = 20, shape = 3, scale = 2)
+y = rnorm(n = 20, mean = 10, sd = 2)
+p = cbind(x, y)
+# Create digital convex hull
+d_ch = digital_convex_hull(p, mins=c(0,5), maxs=c(15,15), spacings = c(0.05,0.05))
+# Plot point data and digital convex hull
+image(x=d_ch[[3]][[1]], y=d_ch[[3]][[2]], z=d_ch[[1]], 
+      yaxt="n", xaxt="n", xlab="", ylab = "", col=c("white", "orange"))
+points(p, pch=16, cex=0.75)
+\end{lstlisting}
 
@@
 \centering
-\includegraphics{figures/figure-2/figure-2.png}
+\includegraphics{figures/digital-algorithms/digital-algorithms.png}
 \caption{Two-dimensional examples of digital geometry algorithms currently available in \texttt{compGeometeR}. (a) Convex hull, (b) alpha complex, and (c) alpha shape.}
@@
 
-\section{Code availability}
+\section{Software availability}
 
@@
 Installation instructions can be found at the GitHub repository \url{https://github.com/manaakiwhenua/compGeometeR}.
+
+We are also developing a cookbook of examples of \texttt{compGeometeR} use via the GitHub repository wiki \url{https://github.com/manaakiwhenua/compGeometeR/wiki}.
 

### 2021-03-10 paper/figures/digital-algorithms/digital-algorithms.R

@@ -14,7 +14,7 @@ p = cbind(x, y)
 
 #-------------------------------------------------------------------------------
 
-png("figure-2.png", width = 15, height = 5, res=300, units = "cm", pointsize = 10)
+png("digital-algorithms.png", width = 15, height = 5, res=300, units = "cm", pointsize = 10)
 par(mfrow=c(1,3), oma=c(0,0,0,0), mar=c(0.5,0.5,1.5,0.5))
 cols = c("white", "slateblue", "firebrick", "lightseagreen", "purple", "darkorchid", "dodgerblue",
          "goldenrod", "grey", "deeppink", "turquoise", "yellowgreen", "yellow", "limegreen",

### 2021-03-10 paper/figures/discrete-algorithms/discrete-algorithms.R

@@ -14,7 +14,7 @@ p = cbind(x, y)
 
 #-------------------------------------------------------------------------------
 
-png("figure-1.png", width = 10, height = 10, res=300, units = "cm", pointsize = 10)
+png("discrete-algorithms.png", width = 10, height = 10, res=300, units = "cm", pointsize = 10)
 par(mfrow=c(2,2), oma=c(0,0,0,0), mar=c(0.5,0.5,1.5,0.5))
 
 # Convex hull

### 2021-03-10 R/digital-alpha-shape.R
Updated digital_alpha_shape function

@@ -44,22 +44,13 @@
 #' @export
 digital_alpha_shape <- function(points=NULL, alpha=Inf, mins, maxs, spacings) {
 
-  # Create the discrete alpha complex
-  ac <- alpha_complex(points = p, alpha = alpha)
-  # Generate a grid of coordinates
-  grid <- grid_coordinates(mins, maxs, spacings)
-  # Check if the grid coordinates are in a simplex
-  m <- find_simplex(ac, grid[[1]])
-  m[m >= 1] <- 1
-  # Get the grid length of each dimension
-  dim_n <- c()
-  for (dim in grid[[2]]) {
-    dim_n <- c(dim_n, length(dim))
-  }
-  # Create an array of the results
-  as_array <- array(m, dim=dim_n)
+  # Create the digital alpha complex
+  d_ac = digital_alpha_complex(points, alpha, mins, maxs, spacings)
+  # Identify grid coordinaes in any simplex
+  d_ac[[1]][d_ac[[1]] >=1] = 1
+  d_ac[[2]][,3][d_ac[[2]][,3] >= 1] = 1
   
-  return(list(as_array, cbind(grid[[1]], m), grid[[2]]))
+  return(d_ac)
   
 }
 

### 2021-03-10 paper/main.tex
Updated paper

@@
 
-Computational geometry seeks to develop algorithms and data structures for geometric objects, and the results are widely applied across numerous scientific domains \citep{de-berg-2008}.
-Given the importance of computational geometry it is no surprise that within the \texttt{R} computing ecosystem there are a variety of options for applying various computational geometry algorithms.
+Geometry is an ancient field of study in which geometer examines the distance, shape, size, and position of objects in space \citep{gowers-2003}.
+Computational geometry is a more recent field of study that emerged alongside technological developments in computing.
+Computational geometry seeks to develop efficient algorithms and data structures for geometric objects, and these algorithms and data structures have been found to be widely applicable across numerous scientific domains \citep{de-berg-2008}.
+
+Given the importance of computational geometry it is no surprise that within the \texttt{R} \citep{r-core-team-2019} computing ecosystem there are a variety of options for applying various computational geometry algorithms.
 However, these packages often work in specific numbers of dimensions, have incompatible data structures, and include additional non-computational geometry functionality that can be domain specific.
 
-Our objective in developing the \texttt{compGeometeR} package is to produce a systematic implementation of some of the most commonly used computational geometry algorithms across various of branches of geometry so that they can be easily integrated into domain specific scientific workflows.
+Our objective in developing the \texttt{compGeometeR} (a computational geometer using \texttt{R}) package is to produce a systematic implementation of some of the most commonly used computational geometry algorithms across various of branches of geometry so that they can be easily integrated into domain specific scientific workflows.
+Our hope is that by producing \texttt{compGeometeR} we can encourage the use of geometry in \texttt{R} for scientific study.
 
@@
 
-The functions of \texttt{compGeometeR} interconnect and build upon one another, but it is important to recognise that virtually all of the \texttt{compGeometeR} functions are  dependent on computational geometry foundations provided by the \texttt{Qhull} C++ interface software \citep{barber-1996}.
+The functions of \texttt{compGeometeR} interconnect and build upon one another, but it is important to recognise that virtually all of the \texttt{compGeometeR} functions are  dependent on computational geometry foundations provided by the \texttt{Qhull} C++ interface software \citep[\url{http://www.qhull.org/}]{barber-1996}.
 
@@
 
-Discrete geometry studies the shapes and interactions between sets of discrete objects such as points, lines, circles, and polygons in 2-dimensional space.
+Discrete geometry studies the shapes and interactions between sets of discrete objects such as points, lines, circles, and polygons in 2-dimensional Euclidean space -- and the hyperdimensional equivalents of these objects in $n$-dimensional Euclidean space.
 All of the discrete computational geometry algorithms in \texttt{compGeometeR} take as an input a set of points $P$ in $n$-dimensional Euclidean space $\mathbb{R}^n$.

### 2021-03-15 paper/main.tex
updated paper

@@
 Computational geometry is a more recent field of study that emerged alongside technological developments in computing.
-Computational geometry seeks to develop efficient algorithms and data structures for geometric objects, and these algorithms and data structures have been found to be widely applicable across numerous scientific domains \citep{de-berg-2008}.
+The two main area of computation geometry are algorithmic and Numerical geometry.
+The algorithmic geometry seeks to develop efficient algorithms and data structures for geometric objects, and these algorithms and data structures have been found to be widely applicable across numerous scientific domains \citep{de-berg-2008}, while the numerical geometry is aim to represent real-world objects which are suitable for computer computations \citep{kimmel2012numerical}.
+The algorithmic geometry is made up of the discrete and digital geometry.
+The discrete algorithmic geometry study the shape and methods of discrete geometric objects, while the digital algorithmic geometry deals with discrete sets  by studying the geometric properties of a grid (or lattice) of points \citep{boissonnat1998algorithmic}.
 
-Given the importance of computational geometry it is no surprise that within the \texttt{R} \citep{r-core-team-2019} computing ecosystem there are a variety of options for applying various computational geometry algorithms.
+Given the importance of computational algorithmic geometry it is no surprise that within the \texttt{R} \citep{r-core-team-2019} computing ecosystem there are a variety of options for applying various computational geometry algorithms.
 However, these packages often work in specific numbers of dimensions, have incompatible data structures, and include additional non-computational geometry functionality that can be domain specific.
+Also, these packages contains only one type of computational geometry algorithmic, which is the discreet geometry.
+The digital part of the computational geometry algorithmic are usually not included  and it is up to the user to further derived the needed data from the output of the discreet geometry (see Table \ref{tab:r-options}).
 
-Our objective in developing the \texttt{compGeometeR} (a computational geometer using \texttt{R}) package is to produce a systematic implementation of some of the most commonly used computational geometry algorithms across various of branches of geometry so that they can be easily integrated into domain specific scientific workflows.
+Our objective in developing the \texttt{compGeometeR} (a computational geometer using \texttt{R}) package is to produce a systematic implementation of some of the most commonly used computational geometry algorithms (Discrete and Digital) across various of branches of geometry so that they can be easily integrated into domain specific scientific workflows.
 Our hope is that by producing \texttt{compGeometeR} we can encourage the use of geometry in \texttt{R} for scientific study.
@@
 
-There are a variety of computational geometry algorithms available in \texttt{compGeometeR}, some of which are replicated in other \texttt{R} packages.
+There are a variety of computational geometry algorithms available in \texttt{compGeometeR}, some of which are also available in other \texttt{R} packages.
 By documenting these differences (Table \ref{tab:r-options}) we hope to help potential users to assess if \texttt{compGeometeR} is the best option for their needs as there may be other options that are more suitable.
@@
 The functions of \texttt{compGeometeR} interconnect and build upon one another, but it is important to recognise that virtually all of the \texttt{compGeometeR} functions are  dependent on computational geometry foundations provided by the \texttt{Qhull} C++ interface software \citep[\url{http://www.qhull.org/}]{barber-1996}.
+Next, we introduce the discrete and digital algorithms included in  \texttt{compGeometeR}
 
@@
   Delaunay triangulation	& $n$ &   &  2  &   & 2 & $n$ & 2 & 2 \\
-  Voronoi diagram			&     &   &  2  &   & 2 & $n$ & 2 & 2 \\
+  Voronoi diagram			&     &   &  2  &   & 2 &  & 2 & 2 \\
 
@@
 \centering
-\includegraphics[width=15cm]{figures/software-structure/software-structure.png}
+\includegraphics[width=15cm]{figures/software-structure/software-structure_arrow_to_be_remove.png}
 \caption{Connections and dependencies of the \texttt{compGeometeR} functions.}
@@
 Digital geometry is a branch of geometry that studies the geometric properties of a grid (or lattice) of points in Euclidean space, and usually involves the digitisation of discrete lines and regions \citep{rosenfeld-1989}.
-Digitisation occurs by representing Euclidean space $\mathbb{R}^n$ as a a rectangular orthogonal grid $\mathbb{G}^n$.
+Digitisation occurs by representing Euclidean space $\mathbb{R}^n$ as a  rectangular orthogonal grid $\mathbb{G}^n$.
 The elements of $\mathbb{G}^2$ are called pixels, and the elements of $\mathbb{G}^3$ are called voxels, and each element in $\mathbb{G}^n$ has a grid coordinate for its centre \citep{klette-2004} and a value that denotes if the element belongs to a digitised geometric object, or when required, which part of the digitised geometric object.
@@
 
-The code required for digital versions of the discrete algorithms is no more complicated, and simply requires addtional parameters to define the extent and resolution of $\mathbb{G}^n$.
+The code required for digital versions of the discrete algorithms is no more complicated, and simply requires additional parameters to define the extent and resolution of $\mathbb{G}^n$.
 For example, the code required for a digital convex hull (Listing \ref{code:digital-convex-hull-code}) does not differ much from that required for a discrete convex hull (Listing \ref{code:discrete-convex-hull-code}).
@@
 The Voronoi diagram \citep{voronoi-1908, okabe-2000} partitions $n$-dimensional space into a set of polytopes for which each polytope delineates the region of $n$-dimensional space that is closest to each point in $P$.
-The Voronoi diagram would be an obvious addition for the discrete geometry algorithms given its wide implementation in other R geometry packages (Table \ref{tab:r-options}).
+The Voronoi diagram would be an obvious addition for the discrete geometry algorithms given its lack of implementation in other R geometry packages (see Table \ref{tab:r-options}).
 

### 2021-03-18 paper/main.tex
Revised paper

--- 2021-03-10 paper/main.tex
@@
 
-Computational geometry algorithms and data structures are widely applied across numerous scientific domains, and there a variety of \texttt{R} packages that apply computational geometry.
-However, these packages often work in specific numbers of dimensions, have incompatible data structures, and include additional non-computational geometry functionality that can be domain specific.
-Our objective in developing the \texttt{compGeometeR} package is to produce a systematic implementation of the most commonly used computational geometry algorithms so that they can be easily integrated into domain specific scientific workflows.
-We briefly explain the algorithms available in \texttt{compGeometeR}, and identify priorities for future development.
+Computational geometry algorithms and data structures are widely applied across numerous scientific domains, and there a variety of \texttt{R} packages that implement computational geometry functionality.
+However, these packages often work in specific numbers of dimensions, do not have directly compatible data structures, and include additional non-computational geometry functionality that can be domain specific.
+Our objective in developing the \texttt{compGeometeR} package is to implement in a generic and consistent framework the most commonly used combinatorial computational geometry algorithms so that they can be easily combined and integrated into domain specific scientific workflows.
+We briefly explain the discrete and digital combinatorial computational geometry algorithms available in \texttt{compGeometeR}, and identify priorities for future development.
 
@@
 
-Computational geometry seeks to develop algorithms and data structures for geometric objects, and the results are widely applied across numerous scientific domains \citep{de-berg-2008}.
-Given the importance of computational geometry it is no surprise that within the \texttt{R} computing ecosystem there are a variety of options for applying various computational geometry algorithms.
-However, these packages often work in specific numbers of dimensions, have incompatible data structures, and include additional non-computational geometry functionality that can be domain specific.
+Geometry is an ancient form of mathematics that enables a geometer to study the distance, shape, size, and position of objects in space \citep{gowers-2003}.
+Computational geometry is a more recent field of study that emerged alongside technological developments in computing and seeks to develop efficient algorithms and data structures for geometric objects.
+The two main areas of computational geometry are numerical computational geometry that considers continuous geometric objects \citep{kimmel-2012} and combinatorial (or algorithmic) computational geometry that considers discrete geometric objects \citep{boissonnat-1998, de-berg-2008}.
 
-Our objective in developing the \texttt{compGeometeR} package is to produce a systematic implementation of some of the most commonly used computational geometry algorithms across various of branches of geometry so that they can be easily integrated into domain specific scientific workflows.
+Our focus here is on combinatorial computational geometry, as these algorithms and data structures have been found to be widely applicable across numerous scientific domains \citep{de-berg-2008}.
+Given the importance of combinatorial computational geometry it is no surprise that within the \texttt{R} \citep{r-core-team-2019} computing ecosystem there are a variety of options for applying various combinatorial computational geometry algorithms.
+However, these algorithms are currently spread across multiple somewhat incompatible packages that often work in specific numbers of dimensions (Table \ref{tab:r-options}) and some include additional non-computational geometry functionality that can be domain specific.
+Also, these packages contain only one type of combinatorial computational geometry algorithm, those based on discrete objects such as points, lines, polygons.
+But other forms of combinatorial computational geometry such as digital geometry that studies the geometric properties of a grid (or lattice) of points \citep{rosenfeld-1989, klette-2004} would be a useful addition to the \texttt{R} language.
 
-\section{Algorithms}
+Our objective in developing the \texttt{compGeometeR} (a computational geometer using \texttt{R}!) package is to produce a systematic implementation of some of the more commonly used discrete and digital combinatorial computational geometry algorithms so that they can be easily integrated into domain specific scientific workflows.
+Our hope is that by producing \texttt{compGeometeR} we can encourage the use of geometry in \texttt{R} for scientific study.
 
-There are a variety of computational geometry algorithms available in \texttt{compGeometeR}, some of which are replicated in other \texttt{R} packages.
-By documenting these differences (Table \ref{tab:r-options}) we hope to help potential users to assess if \texttt{compGeometeR} is the best option for their needs as there may be other options that are more suitable.
-
-The functions of \texttt{compGeometeR} interconnect and build upon one another, but it is important to recognise that virtually all of the \texttt{compGeometeR} functions are  dependent on computational geometry foundations provided by the \texttt{Qhull} C++ interface software \citep{barber-1996}.
-
-% The following table uses formatting fucntions of the booktabs package: https://texdoc.org/serve/booktabs/0
+% The following table uses formatting functions of the booktabs package: https://texdoc.org/serve/booktabs/0
 \begin{table}
@@
   \cmidrule{1-9} 
-  Discrete geometry		&     &   &     &   &   &     &   &   \\
+  Discrete algorithms       &     &   &     &   &   &     &   &   \\
   \cmidrule{1-9} 
@@
   Delaunay triangulation	& $n$ &   &  2  &   & 2 & $n$ & 2 & 2 \\
-  Voronoi diagram			&     &   &  2  &   & 2 & $n$ & 2 & 2 \\
+  Voronoi diagram			&     &   &  2  &   & 2 &     & 2 & 2 \\
 
   \cmidrule{1-9} 
-  Digital geometry			&     &   &     &   &   &     &   &   \\
+  Digital algorithms     	&     &   &     &   &   &     &   &   \\
   \cmidrule{1-9} 
@@
 
-\begin{figure}[ht]
+\section{Algorithms}
+
+The algorithms of \texttt{compGeometeR} that are described in the following sections can be grouped into discrete algorithms and digital algorithms (Table \ref{tab:r-options}), and the functions of \texttt{compGeometeR} that implement both kinds of algorithms have been specifically designed to easily interconnect and build upon one another (Figure \ref{fig:software-structure}).
+While some of the \texttt{compGeometeR} code is new, it is important to recognise that virtually all of the \texttt{compGeometeR} functions are  dependent on computational geometry foundations provided by the \texttt{Qhull} C++ interface software \citep[\url{http://www.qhull.org/}]{barber-1996} (Figure \ref{fig:software-structure}).
+
+It is also important to note that while \texttt{compGeometeR}'s digital algorithms are unique to \texttt{R}, some of the discrete algorithms available in \texttt{compGeometeR} are also available in other \texttt{R} packages (Table \ref{tab:r-options}).
+By documenting these similarities and differences we hope to help potential users to assess if \texttt{compGeometeR} is the best option for their needs as there may be other options that are more suitable.
+
+\begin{figure}[t]
 \centering
@@
 
-\subsection{Discrete geometry}
-% see: https://en.wikipedia.org/wiki/Discrete_geometry
+\subsection{Discrete algorithms}
 
-Discrete geometry studies the shapes and interactions between sets of discrete objects such as points, lines, circles, and polygons in 2-dimensional space.
-All of the discrete computational geometry algorithms in \texttt{compGeometeR} take as an input a set of points $P$ in $n$-dimensional Euclidean space $\mathbb{R}^n$.
+The discrete combinatorial computational geometry algorithms calculate the shapes and interactions between sets of discrete objects such as points, lines, circles, and polygons in 2-dimensional Euclidean space -- and the hyperdimensional equivalents of these objects in $n$-dimensional Euclidean space.
+All of the discrete algorithms in \texttt{compGeometeR} take as an input a set of points $P$ in $n$-dimensional Euclidean space $\mathbb{R}^n$.
 
@@
 
-The Delaunay triangulation \citep{delaunay-1934} produces a set of simplices (triangles in 2-dimensions or tetrahedrons in 3-dimensions) for which no point in $P$ is inside the the circumhypersphere (a circle in 2-dimensions or a sphere in 3-dimensions) of each simplex (Figure \ref{fig:discrete-algorithms}c).
+The Delaunay triangulation \citep{delaunay-1934} produces a set of simplices (triangles in 2-dimensions or tetrahedrons in 3-dimensions) for which no point in $P$ is inside the circumhypersphere (a circle in 2-dimensions or a sphere in 3-dimensions) of each simplex (Figure \ref{fig:discrete-algorithms}c).
 
@@
 
-\subsection{Digital geometry}
-% see: https://en.wikipedia.org/wiki/Digital_geometry
+\subsection{Digital algorithms}
 
-Digital geometry is a branch of geometry that studies the geometric properties of a grid (or lattice) of points in Euclidean space, and usually involves the digitisation of discrete lines and regions \citep{rosenfeld-1989}.
-Digitisation occurs by representing Euclidean space $\mathbb{R}^n$ as a a rectangular orthogonal grid $\mathbb{G}^n$.
+The digital combinatorial computational geometry algorithms calculate the geometric properties of a grid (or lattice) of points in Euclidean space, and usually involves the digitisation of discrete geometric objects \citep{rosenfeld-1989}.
+Digitisation occurs by representing Euclidean space $\mathbb{R}^n$ as a rectangular orthogonal grid $\mathbb{G}^n$.
 The elements of $\mathbb{G}^2$ are called pixels, and the elements of $\mathbb{G}^3$ are called voxels, and each element in $\mathbb{G}^n$ has a grid coordinate for its centre \citep{klette-2004} and a value that denotes if the element belongs to a digitised geometric object, or when required, which part of the digitised geometric object.
@@
 
-The code required for digital versions of the discrete algorithms is no more complicated, and simply requires addtional parameters to define the extent and resolution of $\mathbb{G}^n$.
+The code required for digital versions of the discrete algorithms is only slightly more complex, and simply requires additional parameters to define the extent and resolution of $\mathbb{G}^n$.
 For example, the code required for a digital convex hull (Listing \ref{code:digital-convex-hull-code}) does not differ much from that required for a discrete convex hull (Listing \ref{code:discrete-convex-hull-code}).
@@
 
-\texttt{compGeometeR} is very much a work in progress, and while there is sufficient functionality for it to be of wide use in the computational sciences, there is some functionality that has been identified as being particularly useful for future development.
+\texttt{compGeometeR} is very much a work in progress, and while there is already sufficient functionality for \texttt{compGeometeR} to be useful in a wide range of computational sciences, there is some functionality that has been identified as being particularly useful for future development.
 
 The Voronoi diagram \citep{voronoi-1908, okabe-2000} partitions $n$-dimensional space into a set of polytopes for which each polytope delineates the region of $n$-dimensional space that is closest to each point in $P$.
-The Voronoi diagram would be an obvious addition for the discrete geometry algorithms given its wide implementation in other R geometry packages (Table \ref{tab:r-options}).
+The Voronoi diagram would be an obvious addition for the discrete geometry algorithms given a 2-dimensional version has been implemented in other \texttt{R} geometry packages but no $n$-dimensional version is available (Table \ref{tab:r-options}).
 
@@
 In some contexts it will be important to recognise that there is uncertainty in the position of points in space and the parameters defining a given algorithm.
-In such situations it may be helpful to adopt fuzzy geometry \citep{rosenfeld-1998} view in which, foe example, memberships of points in space are not crisp being either 0 or 1, but rather are fuzzy on a scale from 0 to 1.
+In such situations it may be helpful to adopt fuzzy geometry \citep{rosenfeld-1998} view in which, for example, memberships of points in space are not crisp being either 0 or 1, but rather are fuzzy on a scale from 0 to 1.
 Creating fuzzy versions of the \texttt{compGeometeR} algorithms could result in very useful functionality where quantifying uncertainty of any computational geometry is important.
@@
 All the algorithms in \texttt{compGeometeR} work in Euclidean space, but computational geometry algorithms can also be usefully applied in spherical or hyperbolic space \citep{gowers-2003}.
-For example, the Fortran \texttt{STRIPACK} software \citep{renka-1997} could be wrapped by R to provide functionality to compute Delaunay triangulations and Voronoi diagrams on the surface of a sphere.
+For example, the Fortran \texttt{STRIPACK} software \citep{renka-1997} could be wrapped by \texttt{R} to provide functionality to compute Delaunay triangulations and Voronoi diagrams on the surface of a sphere.
 
@@
 
-\texttt{compGeometeR} is open source software made available under a General Public License license.
+\texttt{compGeometeR} is open source software made available under a General Public License.
 Installation instructions can be found at the GitHub repository \url{https://github.com/manaakiwhenua/compGeometeR}.

### 2021-03-18 R/digital-alpha-complex.R
Simplified digital function outputs and expanded examples

@@ -19,15 +19,13 @@
 #' @param spacings Vector of length \code{d} listing the grid coordinate spacing 
 #' for each dimension.
 #' 
-#' @return A list of three objects:
+#' @return A list of two objects:
 #' 
 #' \itemize{
 #'   \item A \eqn{d}-dimensional array containing an integer index of the alpha 
 #'   complex \eqn{s} \href{https://en.wikipedia.org/wiki/Simplex}{simplex} that 
 #'   each grid coordinate lies within, or 0 if it lies outside the alpha complex 
 #'   (if any of the test point coordinates contain NA then the output is 0).
-#'   \item A dataframe with \code{d} columns and a row for each grid coordinate 
-#'   - so potentially lots of rows!
 #'   \item A list of length \code{d} that contains the grid coordinates along 
 #'   each dimension.
 #' }
@@ -37,10 +35,15 @@
 #' x <- c(30, 70, 20, 50, 40, 70)
 #' y <- c(35, 80, 70, 50, 60, 20)
 #' p <- data.frame(x, y)
-#' # Create digital alpha complex and plot
+#' # Create digital alpha complex
 #' d_ac <- digital_alpha_complex(points = p, alpha = 20, mins=c(15,15), maxs=c(85,85), spacings=c(0.5,0.5))
 #' cols = c("lightgrey", "orange", "purple", "lightseagreen")
-#' image(x=d_ac[[3]][[1]], y=d_ac[[3]][[2]], z=d_ac[[1]], xlab="x", ylab="y", col = cols)
+#' # Unpack the digital alpha complex components
+#' d_ac_array <- d_ac[[1]]
+#' x_coords <- d_ac[[2]][[1]]
+#' y_coords <- d_ac[[2]][[2]]
+#' # Plot the digital convex hull
+#' image(x=x_coords, y=y_coords, z=d_ac_array, xlab="x", ylab="y", col = cols)
 #' points(p, pch = as.character(seq(nrow(p))))
 #' legend("bottomleft", pch=15, col=cols, legend=sort(unique(c(d_ac[[1]]))), title="Simplex")
 #' points(p, pch = as.character(seq(nrow(p))))
@@ -62,7 +65,7 @@ digital_alpha_complex <- function(points=NULL, alpha=Inf, mins, maxs, spacings) 
   # Create an array of the results
   ac_array <- array(m, dim=dim_n)
   
[…]

### 2021-03-18 R/digital-alpha-shape.R

@@ -19,14 +19,12 @@
 #' @param spacings Vector of length \code{d} listing the grid coordinate spacing 
 #' for each dimension.
 #' 
-#' @return A list of three objects:
+#' @return A list of two objects:
 #' 
 #' \itemize{
 #'   \item A \eqn{d}-dimensional array containing 1 if a grid coordinate lies 
 #'   inside the alpha shape and 0 if it lies outside (if any of the test point 
 #'   coordinates contain NA then the output is 0).
-#'   \item A dataframe with \code{d} columns and a row for each grid coordinate 
-#'   - so potentially lots of rows!
 #'   \item A list of length \code{d} that contains the grid coordinates along 
 #'   each dimension.
 #' }
@@ -36,19 +34,23 @@
 #' x <- c(30, 70, 20, 50, 40, 70)
 #' y <- c(35, 80, 70, 50, 60, 20)
 #' p <- data.frame(x, y)
-#' # Create digital alpha complex and plot
+#' # Create digital alpha shape
 #' d_as <- digital_alpha_shape(points = p, alpha = 20, mins=c(15,15), maxs=c(85,85), spacings=c(0.5,0.5))
-#' image(x=d_as[[3]][[1]], y=d_as[[3]][[2]], z=d_as[[1]], xlab="x", ylab="y")
+#' # Unpack the digital alpha shape components
+#' d_as_array <- d_as[[1]]
+#' x_coords <- d_as[[2]][[1]]
+#' y_coords <- d_as[[2]][[2]]
+#' # Plot the digital alpha shape
+#' image(x=x_coords, y=y_coords, z=d_as_array, xlab="x", ylab="y")
 #' points(p, pch = as.character(seq(nrow(p))))
 #' 
 #' @export
 digital_alpha_shape <- function(points=NULL, alpha=Inf, mins, maxs, spacings) {
 
   # Create the digital alpha complex
-  d_ac = digital_alpha_complex(points, alpha, mins, maxs, spacings)
-  # Identify grid coordinaes in any simplex
-  d_ac[[1]][d_ac[[1]] >=1] = 1
-  d_ac[[2]][,3][d_ac[[2]][,3] >= 1] = 1
+  d_ac <- digital_alpha_complex(points, alpha, mins, maxs, spacings)
+  # Identify grid coordinates in any simplex
+  d_ac[[1]][d_ac[[1]] >=1] <- 1
   
   return(d_ac)
   

### 2021-03-18 R/digital-convex-hull.R

@@ -15,14 +15,12 @@
 #' @param spacings Vector of length \code{d} listing the grid coordinate spacing 
 #' for each dimension.
 #' 
-#' @return A list of three objects:
+#' @return A list of two objects:
 #' 
 #' \itemize{
 #'   \item A \eqn{d}-dimensional array containing 1 if a grid coordinate lies 
 #'   within the hull and 0 if it lies outside the hull (if any of the test point 
 #'   coordinates contain NA then the output is 0).
-#'   \item A dataframe with \code{d} columns and a row for each grid coordinate 
-#'   - so potentially lots of rows!
 #'   \item A list of length \code{d} that contains the grid coordinates along 
 #'   each dimension.
 #' }
@@ -32,9 +30,14 @@
 #' x <- c(30, 70, 20, 50, 40, 70)
 #' y <- c(35, 80, 70, 50, 60, 20)
 #' p <- data.frame(x, y)
-#' # Create digital convex hull and plot
+#' # Create digital convex hull
 #' d_ch <- digital_convex_hull(points = p, mins=c(15,15), maxs=c(85,85), spacings=c(0.5,0.5))
-#' image(x=d_ch[[3]][[1]], y=d_ch[[3]][[2]], z=d_ch[[1]], xlab="x", ylab="y")
+#' # Unpack the digital convex hull components
+#' d_ch_array <- d_ch[[1]]
+#' x_coords <- d_ch[[2]][[1]]
+#' y_coords <- d_ch[[2]][[2]]
+#' # Plot the digital convex hull
+#' image(x=x_coords, y=y_coords, z=d_ch_array, xlab="x", ylab="y")
 #' points(p, pch = as.character(seq(nrow(p))))
 #' 
 #' @export
@@ -54,6 +57,6 @@ digital_convex_hull <- function(points=NULL, mins, maxs, spacings) {
   # Create an array of the results
   ch_array <- array(m, dim=dim_n)
   
-  return(list(ch_array, cbind(grid[[1]], m), grid[[2]]))
+  return(list(ch_array, grid[[2]]))
   
 }

### 2021-03-18 paper/main.tex
Updated code examples in the paper

@@
 
-\begin{lstlisting}[language=R, caption=Example \texttt{R} code to create a discrete convex hull with \texttt{compGeometeR}, label={code:discrete-convex-hull-code}]
+\begin{lstlisting}[language=R, caption=Example \texttt{R} code to create a discrete convex hull (Figure \ref{fig:discrete-algorithms}a) with \texttt{compGeometeR}, label={code:discrete-convex-hull-code}]
 library(compGeometeR)
@@
 
-\begin{lstlisting}[language=R, caption=Example \texttt{R} code to create a digital convex hull with \texttt{compGeometeR}, label={code:digital-convex-hull-code}]
+\begin{lstlisting}[language=R, caption=Example \texttt{R} code to create a digital convex hull (Figure \ref{fig:digital-algorithms}a) with \texttt{compGeometeR}, label={code:digital-convex-hull-code}]
 library(compGeometeR)
@@
 d_ch = digital_convex_hull(p, mins=c(0,5), maxs=c(15,15), spacings = c(0.05,0.05))
-# Plot point data and digital convex hull
-image(x=d_ch[[3]][[1]], y=d_ch[[3]][[2]], z=d_ch[[1]], 
+# Unpack the digital convex hull components and plot
+d_ch_array <- d_ch[[1]]
+x_coords <- d_ch[[2]][[1]]
+y_coords <- d_ch[[2]][[2]]
+image(x=x_coords, y=y_coords, z=d_ch_array, 
       yaxt="n", xaxt="n", xlab="", ylab = "", col=c("white", "orange"))

### 2021-03-18 paper/figures/digital-algorithms/digital-algorithms.R

@@ -22,21 +22,25 @@ cols = c("white", "slateblue", "firebrick", "lightseagreen", "purple", "darkorch
 
 # Digital convex hull
 d_ch = digital_convex_hull(p, mins=c(0,5), maxs=c(15,15), spacings = c(0.05,0.05))
-image(x=d_ch[[3]][[1]], y=d_ch[[3]][[2]], z=d_ch[[1]], 
+# Unpack the digital convex hull components and plot
+d_ch_array <- d_ch[[1]]
+x_coords <- d_ch[[2]][[1]]
+y_coords <- d_ch[[2]][[2]]
+image(x=x_coords, y=y_coords, z=d_ch_array, 
       yaxt="n", xaxt="n", xlab="", ylab = "", col=c("white", "orange"))
 points(p, pch=16, cex=0.75)
 mtext("(a)", side = 3, line=0.25)
 
 # Digital alpha complex
 d_ac = digital_alpha_complex(p, alpha = 2, mins=c(0,5), maxs=c(15,15), spacings = c(0.05,0.05))
-image(x=d_ac[[3]][[1]], y=d_ac[[3]][[2]], z=d_ac[[1]]+1, 
+image(x=d_ac[[2]][[1]], y=d_ac[[2]][[2]], z=d_ac[[1]]+1, 
       yaxt="n", xaxt="n", xlab="", ylab = "", col=cols)
 points(p, pch=16, cex=0.75)
 mtext("(b)", side = 3, line=0.25)
 
 # Digital alpha shape
 d_as = digital_alpha_shape(p, alpha = 2, mins=c(0,5), maxs=c(15,15), spacings = c(0.05,0.05))
-image(x=d_as[[3]][[1]], y=d_as[[3]][[2]], z=d_as[[1]], 
+image(x=d_as[[2]][[1]], y=d_as[[2]][[2]], z=d_as[[1]], 
       yaxt="n", xaxt="n", xlab="", ylab = "", col=c("white", "orange"))
 points(p, pch=16, cex=0.75)
 mtext("(c)", side = 3, line=0.25)
